### 2016-09-30 outline.rmd
adding stuff on tesco and on the combinatorial problems

# 'Have you forgotten?'

@@

@@

The shift from a handwritten list to 'Have you forgotten?' is perhaps a **development in memory practices.** It reminds me of the game 'I went shopping and I bought ... ' that I play with children sometimes at the dinner table to settle them down. The game is about remembering a well-ordered list of items - a is apple, b is for banana, c is for carrot.  The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items. The longer the game continues, the harder it is to remember since the shopping list grows longer. At the same time, the many repetitions of this list, the fact that each player recites the whole list, helps memorisation. One trick in playing this game is to choose items that are quite unexpected or increasingly random. Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children. The children have better memories for items they know, but adults know more unexpected items. On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children. Balancing between the improbable and memorable ...

@@

> predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events [@Hansen_2015b, 111-112]

@@

>Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015b, 120]

@@

# 1. The context of the presentation - Consumption and shopping
- presentation at an academic event with speakers from industry and government on statistics and operations research in industry. Speakers came from Shell, Disney, GCHQ, and various universities. 
- Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers. [@Humby_1984]  
Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Humby_1984] 
![Humby, 1984]( figure/Humby_young0R_1994.png )
- the problems of scale: Tesco is like Amazon or Netflix in this respect: too many people and too many products

@@

Shreena described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' 
This was a significant change in how the 'Have You Forgotten' list was constructed. 
I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda [@Agrawal_1994].
The `apriori` algorithm finds sets of items that commonly occur in transactions. 
These commonly occurring sets are expressed as association rules. 
For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart might have discovered  -- but didn't -- the fabled association between diapers and beer.  [http://www.dssresources.com/newsletters/66.php]
It is a key algorithm in 'market basket analysis' and attempts to re-structure shopping lists.

@@

The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of data-mining approaches to lists. 
The association between items is expressed in the form of _rules_. These rules, a term derived from decision support literature, define sets of associations between items. 
As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.   
`r inspect(rules[1])` is a stronger association than `r inspect(rules[2])`.
So any recommendation in the 'Have You Forgotten' list should probably put them in that order. 
Association rules and the `apriori` algorithm present social order in terms of sets of common associations. 
This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou. 
Badiou in his account of multiples highlight the contrast between two different ways of grasping a multiple -- in terms of the 'numerical one' and the 'vital One' [@Badiou_2004, 72], between the 'closed' and the 'open' respectively.  
The `apriori` algorithm, might be seen as an attempt to deal with the excess of things in a Tesco supermarket, and the lack of interest or engagement that people have with them by identifying subsets whose membership is stronger by virtue of the other sets of associations that people bring with them.   
Association is stronger when it derives from an excess of different ways of being in sets.  
Badiou's argument is that being part of a set differs from belonging to a set.
An open set arises from when we dwell within a set by 'taking the intersection of two elements.' 
What might this mean in terms of ordering?
If we take the itemset from the `Groceries` dataset [@Hahsler_2006], and then run it through the Tesco online groceries site, each of the 169 items in the dataset yields dozens and sometimes thousands of results from Tesco online.  
Each of the items in the `Groceries` dataset proliferates in the Tesco site, into a new list of branded variations.

@@

Patel observed that the 'Have you forgotten' lists matter to people.
Some people log on and go straight to the checkout, allowing the recommendation system to supply them with the suggestions that they then move to the shopping basket.
This is the online equivalent of using an old shopping list and only adding new items when needed. 
It is also a strategy to avoid having to view too many lists, since the only way Tesco can show people what they might buy is by showing them lists of things (e.g. chicken products), but these lists are too long to work with.  
People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices.

# 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?
- history of data accumulations;
- shift from data mining to personalization?

# 6. Gain: use good features to build models 
- what counts as a good feature? See Domingos on this

@@

# Conclusion

>>>>

# 'Have you forgotten?' Conversion experiences in list worlds

@@

The question: which five items and in what order?

@@

The shift from a handwritten list to 'Have you forgotten?' is perhaps a **development in memory practices.** It reminds me of the game 'I went shopping and I bought ... ' that I play with children sometimes at the dinner table to settle them down. The game is about remembering a well-ordered list of items - 'a' is apple, 'b' is for banana, 'c' is for carrot.  The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items. The longer the game continues, the harder it is to remember since the shopping list grows longer. At the same time, the many repetitions of this list, the fact that each player recites the whole list, helps memorisation. One trick in playing this game is to choose items that are quite unexpected or increasingly random. Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children. The children have better memories for items they know, but adults know more unexpected items. On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children. Balancing between the improbable and memorable ...

@@

> predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events [@Hansen_2015a, 111-112]

@@

>Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015a, 120]

@@

# 1. The context of the presentation - Consumption and shopping
- presentation at an academic event with speakers from industry and government on statistics and operations research in industry. Speakers came from Shell, Disney, GCHQ, and various universities.  Because I've done a lot of work on machine learning and data science more generally, I found the sessions both familiar and occasionally exciting. 
- Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco. I found her presentation on shopping lists to be very practical since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways. 
Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers. [@Humby_1984]  
Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Humby_1984] 
![Humby, 1984]( figure/2_humby_young_org_1994.png)
- the problems of scale: Tesco is like Amazon or Netflix in this respect: too many people and too many products

@@

Shreena described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' 
This was a significant change in how the 'Have You Forgotten' list was constructed from one kind of model to another. For me, the shift in models matters since it amounts to a much more probabilistic structuring of lists.
I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda [@Agrawal_1994].
The `apriori` algorithm finds sets of items that commonly occur in transactions. 
These commonly occurring sets are expressed as association rules. 
For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart might have discovered  -- but didn't -- the fabled association between diapers and beer.  [http://www.dssresources.com/newsletters/66.php]
It is a key algorithm in 'market basket analysis' and attempts to re-structure shopping lists.

@@

The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of data-mining approaches to lists. 
The association between items is expressed in the form of _rules_. These rules, a term derived from decision support literature, define sets of associations between items. 
As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.   
`r inspect(rules[1])` is a stronger association than `r inspect(rules[2])`.
So any recommendation in the 'Have You Forgotten' list should probably put them in that order. 
Association rules and the `apriori` algorithm present social order in terms of sets of common associations. 
This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou. 
Badiou in his account of multiples highlight the contrast between two different ways of grasping a multiple -- in terms of the 'numerical one' and the 'vital One' [@Badiou_2004, 72], between the 'closed' and the 'open' respectively.  
The `apriori` algorithm, might be seen as an attempt to deal with the excess of things in a Tesco supermarket, and the lack of interest or engagement that people have with them by identifying subsets whose membership is stronger by virtue of the other sets of associations that people bring with them.   
Association is stronger when it derives from an excess of different ways of being in sets.  
Badiou's argument is that being part of a set differs from belonging to a set.
An open set arises from when we dwell within a set by 'taking the intersection of two elements.' 
What might this mean in terms of ordering?
If we take the itemset from the `Groceries` dataset [@Hahsler_2006], and then run it through the Tesco online groceries site, each of the 169 items in the dataset yields dozens and sometimes thousands of results from Tesco online.

```{r tesco, cache=TRUE, }
    library(ggplot2)
    tesco = read.csv('data/tesco_groceries.csv')
    total = sum(tesco$actual, na.rm=TRUE) 
    tesco = na.omit(tesco)
    tesco$actual = as.integer(tesco$actual)

    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products')
```

Each of the items in the `Groceries` dataset proliferates in the Tesco site, into a new list of branded variations. The 169 items of the `Groceries` dataset, itself from a local German or Austrian supermarket [@REF_TBA],  expand into roughly `r total` possible items.

@@

Shreena observed that the 'Have you forgotten' lists matter to people, and we can see why in the light of this proliferation of things in the list. A `r total` list of items is quite hard to list let alone read. 
Shreena said that some people log on and go straight to the checkout, allowing the recommendation system to supply them with the suggestions that they then move to the shopping basket.
This is the online equivalent of using an old shopping list and only adding new items when needed. 
It is also a strategy to avoid having to view too many lists, since the only way Tesco can show people what they might buy is by showing them lists of things (e.g. chicken products), but these lists are often too long to work with. For instance, Tesco has more than 1000 chicken products on its site.   
People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices. 
But it also suggests a transformation in the way that lists exist in the world.

# 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?

Shreena described the data they on grocery purchases. 'We have 52 weeks of purchase data for each customer.' 
We know that the construction of lists has been transformed by recommendation systems.
These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent.
In terms of the ways that lists order the world, the amount of data is not so amount as the way in which things in the world take on a list-like existence. 
I want to dwell on this idea for a while. We have seen two different ways in which lists might be constructed in a recommender system. 
The first uses history and existence of previous lists, including the customer's own list, to populate a new list. 
The second uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list. 
The 'relevancy model,' however, tries to do something different. 
The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
With  `r total` products (actually Shreena mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
If a typical shopping list has 20 items, then there are `r choose(total, 20)` possible combinations.
Some of this vast number of possible shopping lists is highly unlikely in terms of shopping lists and histories of shopping, but potentially highly interesting and value as recommendation.   
In some ways, this resembles the Badiouean ontology of the multiple a HERE

> This means, to put it slightly differently, that surrounding any delimited predictive system is a larger field of data—­what I elsewhere call a “surplus of sensibility”—­that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates. (Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy—­the data constituting its real potentiality—­always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015, 126-7]

- history of data accumulations;

- shift from data mining to personalization? 
- use some of the Varian, Linden, Resnick, Schaffer literature

# 6. Gain: use good features to build models 
- what counts as a good feature? See Domingos on this
- move from Dunnhumby's ClubCard focus to much more general world of predictive models (e.g. Facebook use a million models, etc.)

@@

# Conclusion
- lists have complex temporal structure
- listing practices very closely entangled with problems of social order
- lists undergoing a 'conversion' experience as they become more probabilistic
